---
title: "bodyfat"
author: "Giordano Vitale"
date: "2023-07-13"
output: html_document
---


The variables are:

- Density determined from underwater weighing
- Percent body fat from Siri's (1956) equation
- Age (years)
- Weight (lbs)
- Height (inches)
- Neck circumference (cm)
- Chest circumference (cm)
- Abdomen circumference (cm)
- Hip circumference (cm)
- Thigh circumference (cm)
- Knee circumference (cm)
- Ankle circumference (cm)
- Biceps (extended) circumference (cm)
- Forearm circumference (cm)
- Wrist circumference (cm)


# 0) IMPORT LIBRARIES AND DATA
```{r}
bodyfat <- read.csv("C:/Users/giord/OneDrive - Università degli Studi di Milano/1st year/Quarter 2/Statistical Learning/BodyFat Prediction/bodyfat.csv")
```

Convert Weight from lbs to KG and convert Height from inches to CMs.
```{r}
bodyfat$Weight <- bodyfat$Weight/2.2
bodyfat$Height <- bodyfat$Height*2.54
```


```{r}
library(dplyr)
library(purrr)
library(ggplot2)
library(car)
library(corrplot)
library(e1071)
library(MASS)
```

# 1) DATA PREPROCESSING

```{r}
head(bodyfat)[1:3,]
```


```{r}
dim(bodyfat)
```


Check the datatype for each column.
```{r}
map(bodyfat, class)
```

Check for null values.
```{r}
# colSums(is.na(bodyfat)) > 0
sum(is.na(bodyfat))
```

Check for duplicates.
```{r}
sum(duplicated(bodyfat))
```

Check for infinite values.
```{r}
apply(bodyfat, 2, function(x) any(is.infinite(x)))
```

```{r}
sapply(bodyfat, function(x) n_distinct(x, na.rm = TRUE))
```



# 2) EXPLORATORY DATA ANALYSIS

```{r}
summary(bodyfat)
```

Create a function to build histogram plots with normal line fitted.
```{r}
histo_plot<- function (data, col, bins = 30) {
  ggplot(data, aes(x = .data[[col]])) + 
  geom_histogram(aes(y = after_stat(density)),
                 bins = bins,
                 fill = "lightblue",
                 color = "black") + 
  geom_density(kernel = "gaussian",
               color = "red",
               size = 1.2) + 
  labs(x = col, y= "Density") +
  ggtitle(paste("Histogram with density line -", col)) +
  theme(plot.title = element_text(hjust = 0.5))
  }
```


Apply the function to each column.
```{r}
lapply(colnames(bodyfat), function(col) histo_plot(bodyfat, col, bins=16))
```


```{r}
# library(e1071)

skewness <- lapply(colnames(bodyfat), function(col) skewness(bodyfat[[col]]))

variable_names <- names(bodyfat)

skewness <- unlist(skewness)
skewness <- round(skewness, digits = 2)

skewness_df <- data.frame(Variable = variable_names, Skewness = skewness)
skewness_df
```


Check if BodyFat is normally distributed
```{r}
shapiro.test(bodyfat$BodyFat)
```
From the output, the p-value > 0.05 implying that the distribution of the data are not significantly different from normal distribution. In other words, we can assume the normality.

Do it for all the variables:
```{r}
# compute shapiro test for each column of the data set
shapiro_results <- lapply(bodyfat, function(col) shapiro.test(col))

# Extract variable names and p-values from the test results
variable_names <- names(bodyfat)

# compute the p values
p_values <- round(sapply(shapiro_results, function(result) result$p.value), 
                  digits = 2)

# create the dataframe containing results 
shapiro_results_df <- data.frame(Variable = variable_names, P_Value = p_values)
row.names(shapiro_results_df) <- NULL

# add a column that returns Yes if normal, No otherwise
shapiro_results_df$Normal <- ifelse(shapiro_results_df$P_Value > 0.05, "Yes", "No")
shapiro_results_df
```



Take a look at the outlier in Height column.
```{r}
bodyfat[bodyfat$Height < 100, ]
```


Output box plots for each column of the data frame.
```{r}
for (col in colnames(bodyfat)) {
  boxplot(bodyfat[[col]], main = paste("Boxplot - ", col))
}
```

Produce Q-Q plots with confidence intervals.
```{r}
for (col in colnames(bodyfat)) {
  qqPlot(bodyfat[[col]], main = paste("Q-Q plot - ", col))
}
```


Scatter plots of the dependent variable as a function of the i-th independent one, individually.
```{r}
scatter_plot <- function(x_var, y_var) {
  ggplot(bodyfat, aes(x = !!sym(x_var), y = !!sym(y_var))) +
    geom_point() +
    labs(x = x_var, y = y_var) +
    ggtitle(paste("Scatter Plot of", y_var, "vs.", x_var)) +
    theme(plot.margin = unit(c(1,1,1,1), "cm"),      # Set plot margins
         aspect.ratio = 0.5,                            # Set aspect ratio (square)
        plot.background = element_rect(fill = "white"))  # Set plot background color
}
```

```{r}
explanatory_variables <- names(bodyfat)[3:15]

# Create scatter plots for each explanatory variable
scatter_plots_list <- lapply(explanatory_variables, function(var) scatter_plot(var, "BodyFat"))
scatter_plots_list
```


Compute correlation matrix.
```{r}
correlation_matrix <- cor(bodyfat)
round(correlation_matrix, 2)
```

```{r}
corrplot(correlation_matrix, method = "number", 
         number.cex = 0.6, 
         tl.col = "black", 
         type = "lower", 
         tl.srt = 45)
```



Plot it.
```{r}
corrplot(correlation_matrix, method = "circle", 
         type = "lower",
         order = "hclust",
         tl.col = "black", 
         tl.srt = 45)
```


# 3) FEATURE ENGINEERING
Remove Density.
```{r}
bodyfat <- bodyfat %>%
  select(-Density)
```

Create ACratio and HTratio
```{r}
bodyfat$ACratio <- bodyfat$Abdomen / bodyfat$Chest
bodyfat$HTratio <- bodyfat$Hip / bodyfat$Thigh
```

Remove Abdomen, Chest, Hip, Thigh since we created new variables. 
Why? Because of high collinearity.
```{r}
bodyfat <- bodyfat %>% 
  select(-Abdomen, -Chest, -Hip, -Thigh)
```

```{r}
head(bodyfat)
```

# ?? VIF ??
```{r}
simple_model <- lm(BodyFat ~ ., bodyfat)
```

```{r}
vif_values <- vif(simple_model)

barplot(vif_values, main = "VIF" , horiz = "TRUE")
abline(v = 5, lwd = 3, lty = 2)
```



```{r}
correlation_matrix2 <- cor(bodyfat)
round(correlation_matrix2, 2)
```

```{r}
corrplot(correlation_matrix2, method = "number", 
         number.cex = 0.6, 
         tl.col = "black", 
         type = "lower", 
         tl.srt = 45)
```





# ++++++++++++ REGRESSION ANALYSIS ++++++++++++

Apply the train-test split.
```{r}
sample <- sample(c(TRUE, FALSE), 
                 nrow(bodyfat), 
                 replace=TRUE, 
                 prob=c(0.75,0.25))
train  <- bodyfat[sample, ]
test   <- bodyfat[!sample, ]
```


### A) Linear Regression ###

- LINEAR REGRESSION USING PRINCIPAL COMPONENTS [?]

- LINEAR REGRESSION USING ALL THE VARIABLES
```{r}
# Train the model
complete_lm <- lm(BodyFat ~ ., data = train)

# summary of the model
summary(complete_lm)

# see Salini's additional code in file 1
```
Compute measure of goodness.
```{r}
# R Squared
summary(complete_lm)$r.squared

# Adjusted R Squared
summary(complete_lm)$adj.r.squared

# BIC
BIC(complete_lm)

#AIC 
AIC(complete_lm)
```

- ROBUST REGRESSION USING ALL THE VARIABLES
```{r}
# Train the model
robust_lm <- rlm(BodyFat ~ ., data = train, 
                 psi = psi.hampel,
                 init = "lts")

# Summary
summary(robust_lm)

# BIC
BIC(robust_lm)
```




### B) LOGISTIC REGRESSION MODEL ###
```{r}
logistic_regression <- glm(BodyFat ~ ., data = train)
summary(logistic_regression)


# McFadden R squared
McF <- round(with(summary(logistic_regression), 1 - deviance/null.deviance),
             digits = 2)
paste("McFadden R squared for logistic regression is ", McF)
```

### C) - SUBSET SELECTION ###
(http://www.sthda.com/english/articles/37-model-selection-essentials-in-r/155-best-subsets-regression-essentials-in-r/)

```{r}
# train the models
subset_selection <- regsubsets(BodyFat ~ ., data = train, method = "exhaustive", nvmax = 11)

# summary
subset_selection_summary <- summary(subset_selection)
subset_selection_summary
```

```{r}
# optimal number of features to include in the model, according to Cp
which.min(subset_selection_summary$cp)
```

```{r}
# visualize the performance of each "best performing" model with n features (n = 1, ..., 11)
plot(subset_selection_summary$cp, 
     xlab = "Number of Variables",
     ylab = "Adj R Squared",
     main = "Adj-R-Sqared for each best performing model")

points(9, 
       subset_selection_summary$cp[9],
       pch = 0,
       col = "red",
       cex = 2)
```

We saw that the best number of variables to include was 9 according to the Cp measurement.
What about AdjustedRSquared and BIC?
```{r}
data.frame(
  Adj.R2 = which.max(subset_selection_summary$adjr2),
  CP = which.min(subset_selection_summary$cp),
  BIC = which.min(subset_selection_summary$bic)
)
```
Show the coefficients of the best model according to Cp and Adj-R.Squared.
```{r}
coef(subset_selection, 9)

subset_selection_summary$which[9, ]
```

Show the coefficients of the best model according to BIC.
```{r}
coef(subset_selection, 6)

subset_selection_summary$which[6, ]
```

__Comment: even if 9 is the best number of features according to 2 goodness measurements, the improvement from 6 to 9 is negligible and for this reason, for higher simplicity, the model with 6 features may be preferred.__

Formally define the model with 9 variables (found through Cp and Adj-R-Sq)
```{r}
subset9 <- lm(BodyFat ~ Age + Weight + Height + 
                       Neck + Knee + Forearm +
                        Wrist + ACratio + HTratio,
              data = train) 

summary(subset9)
```

Formally define the model with 6 variables (found through BIC)
```{r}
subset6 <- lm(BodyFat ~ Age + Weight + Height + 
                        Forearm + Wrist + ACratio,
              data = train) 

summary(subset6)
```














FOR THE CLUSTERING PART, MAYBE.
We can create a new categorical variable that classifies people in 3 categories: over, normal, below.
```{r}

```