```{r}
library(dplyr)
library(purrr)
library(ggplot2)
library(car)
library(corrplot)
```


```{r}
bodyfat <- read.csv("C:/Users/giord/OneDrive - Università degli Studi di Milano/1st year/Quarter 2/Statistical Learning/BodyFat Prediction/bodyfat.csv")
```


# 2) EXPLORATORY DATA ANALYSIS

```{r}
summary(bodyfat)
```

Create a function to build histogram plots with normal line fitted.
```{r}
histo_plot<- function (data, col, bins = 30) {
  ggplot(data, aes(x = .data[[col]])) + 
  geom_histogram(aes(y = after_stat(density)),
                 bins = bins,
                 fill = "lightblue",
                 color = "black") + 
  geom_density(kernel = "gaussian",
               color = "blue") + 
  labs(x = col, y= "Density") +
  ggtitle(paste("Histogram with density line -", col)) +
  theme(plot.title = element_text(hjust = 0.5))
  }
```


Apply the function to each column.
```{r}
lapply(colnames(bodyfat), function(col) histo_plot(bodyfat, col, bins=16))
```

Check if BodyFat is normally distributed
```{r}
shapiro.test(bodyfat$BodyFat)
```
From the output, the p-value > 0.05 implying that the distribution of the data are not significantly different from normal distribution. In other words, we can assume the normality.


Do it for all the variables:
```{r}
# compute shapiro test for each column of the data set
shapiro_results <- lapply(bodyfat, function(col) shapiro.test(col))

# Extract variable names and p-values from the test results
variable_names <- names(bodyfat)

# compute the p values
p_values <- round(sapply(shapiro_results, function(result) result$p.value), 
                  digits = 2)

# create the dataframe containing results 
data.frame(Variable = variable_names, P_Value = p_values)
```


Take a look at the outlier in Height column.
```{r}
bodyfat[bodyfat$Height < 100, ]
```



Output box plots for each column of the data frame.
```{r}
for (col in colnames(bodyfat)) {
  boxplot(bodyfat[[col]], main = paste("Boxplot - ", col))
}
```

Produce Q-Q plots with confidence intervals.
```{r}
for (col in colnames(bodyfat)) {
  qqPlot(bodyfat[[col]], main = paste("Q-Q plot - ", col))
}
```



Scatter plots of the dependent variable as a function of the i-th independent one, individually.
```{r}
scatter_plot <- function(x_var, y_var) {
  ggplot(bodyfat, aes(x = !!sym(x_var), y = !!sym(y_var))) +
    geom_point() +
    labs(x = x_var, y = y_var) +
    ggtitle(paste("Scatter Plot of", y_var, "vs.", x_var)) +
    theme(plot.margin = unit(c(1,1,1,1), "cm"),      # Set plot margins
         aspect.ratio = 0.5,                            # Set aspect ratio (square)
        plot.background = element_rect(fill = "white"))  # Set plot background color
}
```

```{r}
explanatory_variables <- names(bodyfat)[3:15]

# Create scatter plots for each explanatory variable
scatter_plots_list <- lapply(explanatory_variables, function(var) scatter_plot(var, "BodyFat"))
scatter_plots_list
```



Compute correlation matrix.
```{r}
correlation_matrix <- cor(bodyfat)
round(correlation_matrix, 2)
```

```{r}
corrplot(correlation_matrix, method = "circle", type = "lower", order = "hclust",
         tl.col = "black", tl.srt = 45)
```

A better way to assess multi-collinearity is VIF. First, create the model:
```{r}
simple_model <- lm(BodyFat ~ ., bodyfat)

summary(simple_model)
```
apply VIF to the simple model
```{r}
vif_values <- vif(simple_model)

barplot(vif_values, main ="VIF values", horiz = "TRUE")
abline(v = 5, lwd = 3, lty = 2)
```

```{r}
vif_values
```

Clean the dataset: remove Density, create new variables, as follows
```{r}
# remove density 
bodyfat <- bodyfat[-1]
```

create ACratio
```{r}
bodyfat$ACratio <- bodyfat$Abdomen / bodyfat$Chest
bodyfat$HTratio <- bodyfat$Hip / bodyfat$Thigh
```

Remove Abdomen, Chest, Hip, Thigh since we created new variables.
```{r}
bodyfat <- bodyfat %>% 
  select(-Abdomen, -Chest, -Hip, -Thigh)
```

Remove Weight since it has a VIF value of 34!
```{r}
#bodyfat <- bodyfat %>%
 # select(-Weight)
```

```{r}
head(bodyfat)
```

Re-test a new simple model with these new columns.
```{r}
simple_model2 <- lm(BodyFat ~ ., bodyfat)

summary(simple_model2)
```

```{r}
vif_values2 <- vif(simple_model2)

barplot(vif_values2, main ="VIF values 2", horiz = "TRUE")
abline(v = 5, lwd = 3, lty = 2)
```
Now the VIF outputs a more encouraging group of values.
