```{r}
summary(bodyfat)
```

Create a function to build histogram plots with normal line fitted.
```{r}
histo_plot<- function (data, col, bins = 30) {
  ggplot(data, aes(x = .data[[col]])) + 
  geom_histogram(aes(y = after_stat(density)),
                 bins = bins,
                 fill = "lightblue",
                 color = "black") + 
  geom_density(kernel = "gaussian",
               color = "red",
               size = 1.2) + 
  labs(x = col, y= "Density") +
  ggtitle(paste("Histogram with density line -", col)) +
  theme(plot.title = element_text(hjust = 0.5))
  }
```


Apply the function to each column.
```{r}
lapply(colnames(bodyfat), function(col) histo_plot(bodyfat, col, bins=16))
```


```{r}
library(e1071)

skewness <- lapply(colnames(bodyfat), function(col) skewness(bodyfat[[col]]))

variable_names <- names(bodyfat)

skewness <- unlist(skewness)
skewness <- round(skewness, digits = 2)

skewness_df <- data.frame(Variable = variable_names, Skewness = skewness)
skewness_df
```


Check if BodyFat is normally distributed
```{r}
shapiro.test(bodyfat$BodyFat)
```
From the output, the p-value > 0.05 implying that the distribution of the data are not significantly different from normal distribution. In other words, we can assume the normality.

Do it for all the variables:
```{r}
# compute shapiro test for each column of the data set
shapiro_results <- lapply(bodyfat, function(col) shapiro.test(col))

# Extract variable names and p-values from the test results
variable_names <- names(bodyfat)

# compute the p values
p_values <- round(sapply(shapiro_results, function(result) result$p.value), 
                  digits = 2)

# create the dataframe containing results 
shapiro_results_df <- data.frame(Variable = variable_names, P_Value = p_values)
row.names(shapiro_results_df) <- NULL

# add a column that returns Yes if normal, No otherwise
shapiro_results_df$Normal <- ifelse(shapiro_results_df$P_Value > 0.05, "Yes", "No")
shapiro_results_df
```



Take a look at the outlier in Height column.
```{r}
bodyfat[bodyfat$Height < 100, ]
```


Output box plots for each column of the data frame.
```{r}
for (col in colnames(bodyfat)) {
  boxplot(bodyfat[[col]], main = paste("Boxplot - ", col))
}
```

Produce Q-Q plots with confidence intervals.
```{r}
for (col in colnames(bodyfat)) {
  qqPlot(bodyfat[[col]], main = paste("Q-Q plot - ", col))
}
```


Scatter plots of the dependent variable as a function of the i-th independent one, individually.
```{r}
scatter_plot <- function(x_var, y_var) {
  ggplot(bodyfat, aes(x = !!sym(x_var), y = !!sym(y_var))) +
    geom_point() +
    labs(x = x_var, y = y_var) +
    ggtitle(paste("Scatter Plot of", y_var, "vs.", x_var)) +
    theme(plot.margin = unit(c(1,1,1,1), "cm"),      # Set plot margins
         aspect.ratio = 0.5,                            # Set aspect ratio (square)
        plot.background = element_rect(fill = "white"))  # Set plot background color
}
```

```{r}
explanatory_variables <- names(bodyfat)[3:15]

# Create scatter plots for each explanatory variable
scatter_plots_list <- lapply(explanatory_variables, function(var) scatter_plot(var, "BodyFat"))
scatter_plots_list
```


Compute correlation matrix.
```{r}
correlation_matrix <- cor(bodyfat)
round(correlation_matrix, 2)
```

```{r}
corrplot(correlation_matrix, method = "number", 
         number.cex = 0.6, 
         tl.col = "black", 
         type = "lower", 
         tl.srt = 45)
```



Plot it.
```{r}
corrplot(correlation_matrix, method = "circle", 
         type = "lower",
         order = "hclust",
         tl.col = "black", 
         tl.srt = 45)
```


# 3) FEATURE ENGINEERING
Remove Density.
```{r}
bodyfat <- bodyfat %>%
  select(-Density)
```

Create ACratio and HTratio
```{r}
bodyfat$ACratio <- bodyfat$Abdomen / bodyfat$Chest
bodyfat$HTratio <- bodyfat$Hip / bodyfat$Thigh
```

Remove Abdomen, Chest, Hip, Thigh since we created new variables. 
Why? Because of high collinearity.
```{r}
bodyfat <- bodyfat %>% 
  select(-Abdomen, -Chest, -Hip, -Thigh)
```

```{r}
head(bodyfat)
```

# ?? VIF ??
```{r}
simple_model <- lm(BodyFat ~ ., bodyfat)
```

```{r}
vif_values <- vif(simple_model)

barplot(vif_values, main = "VIF" , horiz = "TRUE")
abline(v = 5, lwd = 3, lty = 2)
```



```{r}
correlation_matrix2 <- cor(bodyfat)
round(correlation_matrix2, 2)
```

```{r}
corrplot(correlation_matrix2, method = "number", 
         number.cex = 0.6, 
         tl.col = "black", 
         type = "lower", 
         tl.srt = 45)
```