# 5) MODEL EVALUATION 


### PERFORMANCE OF THE PREDICTIONS ON TEST DATA ###

Linear model with all variables
```{r}
pred_linear <- predict(complete_lm, newdata = test)
mse_linear <- MSE(pred_linear, test$BodyFat)
```

Robust linear model with all variables
```{r}
pred_robust <- predict(robust_lm, newdata = test)
mse_robust <- MSE(pred_robust, test$BodyFat)
```

Logistic Regression
```{r}
pred_log <- predict(logistic_regression, newdata = test)
mse_log <- MSE(pred_log, test$BodyFat)
```

Subset selection
```{r}
pred_subset_lin <- predict(best_subset, newdata = test)
mse_subset_lin <- MSE(pred_subset, test$BodyFat)
```

Polynomial Regression
```{r}
pred_polyn <- predict(polynomial_model, newdata = test)
mse_polyn <- MSE(pred_polyn, test$BodyFat)
```

Subset Polynomial Regression
```{r}
pred_sub_polyn <- predict(subset_polyn, newdata = test)
mse_sub_polyn <- MSE(pred_sub_polyn, test$BodyFat)
```

PLS components
```{r}
pred_pls_comp <- predict(model_pls_components, newdata = test)
mse_pls_comp <- MSE(pred_pls_comp, test$BodyFat)
```

Local Regression
```{r}
pred_loc_reg <- predict(local_regression, newdata = test)
mse_loc_reg <- MSE(pred_loc_reg, test$BodyFat)
```

Pruned Regression Tree
```{r}
pred_pruned <- predict(pruned_tree, newdata = test)
mse_pruned <- MSE(pred_pruned, test$BodyFat)
```

GAM
```{r}
pred_gam <- predict(gam_model, newdata = test)
mse_gam <- MSE(pred_gam, test$BodyFat)
```

SVM
```{r}
pred_svm <- predict(svm_regression, newdata = test)
mse_svm <- MSE(pred_svm, test$BodyFat)
```



Create a data.frame containing BIC and AdjRSquared for each model
```{r}
model_evaluation <- data.frame(
  
  Models = c("Linear All", "Robust", "Logistic", "Subset Linear", 
             "Polynomial All", "Subset Polynomial", 
             "PLS components", "Local Regression",
             "Pruned Regression Tree", "GAM", "SVM"),
  
  BIC = c(BIC_linear, BIC_robust, BIC_log_reg, BIC_sub,
          BIC_polynomial, BIC_sub_polyn, 
          BIC_pls_components, "NA", 
          "NA", BIC_gam, "NA"),
  
  MSE = c(mse_linear, mse_robust, mse_log, mse_subset_lin, 
          mse_polyn, mse_sub_polyn, 
          "NA", mse_loc_reg,
          mse_pruned, mse_gam, mse_svm)
)

# round the values of the BICs
model_evaluation$BIC <- round(as.numeric(model_evaluation$BIC), digits = 2)

# round the values of the BICs
model_evaluation$MSE <- round(as.numeric(model_evaluation$MSE), digits = 2)

model_evaluation
```

```{r}
model_evaluation[which.min(model_evaluation$MSE), ]
```


```{r}
par(mar = c(2,20,2,2))

barplot(height = model_evaluation$MSE, 
        names.arg = model_evaluation$Models,
        col = "darkolivegreen3",
        main = "MSE comparison among all the models",
        xlab = "MSE",
        xlim = c(0,120),
        ylim = c(0.5, length(model_evaluation$Models) + 3),
        axisnames = TRUE,
        las = 1,
        cex.names = 0.7,
        horiz = TRUE
        )

abline(v = min(model_evaluation$MSE), 
       col = "red",
       lty = 2,
       lwd = 3)

text(min(model_evaluation$MSE) + 22, 
     length(model_evaluation$Models) + 3, 
     paste("Min MSE =", min(model_evaluation$MSE)), 
     pos = 2, 
     col = "red")
```


```{r}
par(mar = c(2,20,2,2))

barplot(height = model_evaluation$BIC, 
        names.arg = model_evaluation$Models,
        col = "lightblue2",
        main = "BIC comparison among all the models",
        xlab = "BIC",
        xlim = c(0,1500),
        ylim = c(0.5, length(model_evaluation$Models) + 3),
        axisnames = TRUE,
        las = 1,
        cex.names = 0.7,
        horiz = TRUE
        )

abline(v = min(model_evaluation$BIC), 
       col = "red",
       lty = 2,
       lwd = 3)

text(min(model_evaluation$BIC) + 22, 
     length(model_evaluation$Models) + 3, 
     paste("Min BIC =", min(model_evaluation$BIC)), 
     pos = 2, 
     col = "red")
```
