### PRINCIPAL COMPONENT ANALYSIS ###
Remove the dependent variable BodyFat.
```{r}
library(dplyr)
X <- bodyfat %>% 
  dplyr::select(-BodyFat)

dim(X)
```

Use the R built-in function `prcomp()` to calculate the principal components of the data set.
```{r}
pca_results <- prcomp(X,
                      scale = TRUE,
                      center = TRUE)
summary(pca_results)
```

Variance of the principal components
```{r}
# for each principal component, we take its standard deviation (see table above) and square it, thus we obtain their variance
pc_var <- pca_results$sdev ** 2
pc_var
```

Percentage of variance explained by the components.
```{r}
explained_var <- pc_var / sum(pc_var)
explained_var
```

The first three components explain the 73% of the variance in the original data set.
```{r}
sum(explained_var[1:3])
```

Plot the eigenvalues (= variances) against the number of dimensions
```{r}
library(factoextra)

fviz_eig(pca_results, 
         addlabels = TRUE,
         xlab = "Principal Components",
         ylab = "Eigenvalue/Variance",
         choice = "eigenvalue",
         barfill = "tan1",
         linecolor = "black",
         barcolor = "black",
         main = "Figure 2",
         ncp = 11) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "red") +
  ggtitle("Sree Plot") +
  theme(plot.title = element_text(hjust = 0.5))
```



```{r}
par(mfrow = c(2,1))
plot(explained_var, type = 'o', ylab = 'Explained Variance', 
     xlab = "Principal Component", col = "blue",
     main = "Variance Explained by each component")
plot(cumsum(explained_var), type = "o", ylab = "Cumulative Exp. Var.", 
     xlab = "Principal Component", col = "brown3",
     main = "Cumulative Variance Explained")
```

Save the first 3 components.
```{r}
pc1 <- pca_results$x[, "PC1"]

pc2 <- pca_results$x[, "PC2"]

pc3 <- pca_results$x[, "PC3"]
```


```{r}
biplot(pca_results, scale = 0,
       xlab = "Principal component 1",
       ylab = "Principal component 2")
```

### PC REGRESSION ###
```{r}
X_train <- train %>% 
  dplyr::select(-BodyFat)

dim(X_train)
```

```{r}
pca_results2 <- prcomp(X_train,
                      scale = TRUE,
                      center = TRUE)
summary(pca_results2)
```
```{r}
pc_var2 <- pca_results2$sdev ** 2
pc_var2
```

Percentage of variance explained by the components.
```{r}
explained_var2 <- pc_var2 / sum(pc_var2)
explained_var2
```

The first three components explain the 73% of the variance in the original data set.
```{r}
sum(explained_var2[1:3])
```
Save the first three components.
```{r}
pc1_train <- pca_results2$x[, "PC1"]
pc2_train <- pca_results2$x[, "PC2"]
pc3_train <- pca_results2$x[, "PC3"]
```


```{r}
pc_regression <- lm(BodyFat ~ pc1_train + pc2_train + pc3_train,
                    data = train)

summary(pc_regression)
```

```{r}
# compute the predictions using pc_regression
pred_pc2 <- predict(pc_regression, newdata = test)
mse_pcr <- MSE(pred_pc2, test$BodyFat)
```
```{r}
BIC(pc_regression)
```